Illustrating imputation of missing data using the Titanic dataset

### 2024-07-21 text/21/negative-binomial-regression.Rmd

```{r, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

```{r negative-binomial-regression-01}
library(MASS)
nb1 <- glm.nb(art ~ fem + ment + phd + mar + kid5, data = articles)
summary(nb1)
```

>>>>

```{r negative-binomial-04, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r negative-binomial-05}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r negative-binomial-06}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

```{r negative-binomial-regression-07}
library(MASS)
nb1 <- glm.nb(art ~ fem + ment + phd + mar + kid5, data = articles)
summary(nb1)
```

### 2024-07-21 text/24/titanic-imputation.Rmd

```{r titanic-imputation-02, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- select(raw_data, o)
vnames <- vnames[o]
head(ti0)
```

Here are a few descriptive statistics

```{r titanic-imputation-03, echo=FALSE}
cat(vnames[1], ": ", vlabels[1], "\n")
summary(ti0$Age)
cat(vnames[2], ": ", vlabels[2], "\n")
count(ti0, PClass)
cat(vnames[3], "\n")
count(ti0, Sex)
cat(vnames[4], "\n")
cat(as.yaml(dd$vars$Survived$values), "\n")
count(ti0, Survived)
```

@@

Create mutliply imputed values for age. The default is to use every variable in the dataset other than age to impute the value of age. You don't want to use the name of the passenger, of course, so be sure to drop it before the imputation step.

```{r titanic-imputation-05}
ti1 <- select(ti0, -Name)
ti1$i_female <- as.numeric(ti1$Sex=="female")
ti1 <- select(ti1, -Sex)
ti_mice <- mice(ti1)
```

@@

The key variable here is sexmale, which shows how much different the log odds of survival are between men and women.

>>>>

```{r titanic-imputation-02, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- raw_data
# ti0 <- raw_data[ , o]
# vnames <- vnames[o]
# vlabels <- vlabels[o]
head(ti0)
```

@@

Create multiply imputed values for age. The default is to use every variable in the dataset other than age to impute the value of age. You don't want to use the name of the passenger, of course, so be sure to drop it before the imputation step.

```{r titanic-imputation-05}
ti1 <- ti0
ti1$i_female <- as.numeric(ti1$Sex=="female")
ti1 <- ti1[ , c("PClass", "Age", "Survived", "i_female")]
ti_mice <- mice(ti1)
```

@@

The key variable here is i_female, which shows how much different the log odds of survival are between men and women.
